Weakly Nonlinear MRI: Notes

### 2014-03-17 notes/mri_notes.tex
two more terms
finished nonlinear terms
perturbation expansion

@@

>>>>

\section*{Expansion of Nonlinear Terms}

@@

Now we expand $N^{(u_{1y})} = J(\Psi, u_{1y}) - \frac{1}{4\pi} J(A, B_{1y})$: \\

$= J(\Psi, u_{1y}) + \epsilon \widetilde{J}(\Psi, u_{1y}) - \frac{1}{4\pi}J(A, B_{1y}) \, - \, \frac{1}{4\pi}\epsilon \widetilde{J}(A, B_{1y})$ \\

For the purposes of the perturbation series expansions, we will rewrite $u_{1y} \equiv u$ and $B_{1y} \equiv B$, with the understanding that $u$ and $B$ are the y-components of the velocity and magnetic field, respectively. \\

$N^{(u_{1y})} \to \epsilon^2 J(\Psi_1, u_1) \, - \, \frac{1}{4\pi} \epsilon^2 J(A_1, B_1) \, + \, \epsilon^3 J(\Psi_1, u_2) \, + \, \epsilon^3 J(\Psi_2, u_1) \, + \, \epsilon^3 \widetilde{J}(\Psi_1, u_1) \, - \, \frac{1}{4\pi} \epsilon^3 J(A_1, B_2) \, - \, \frac{1}{4\pi} \epsilon^3 J(A_2, B_1) \, - \, \frac{1}{4\pi}\epsilon^3 \widetilde{J}(A_1, B_1) \, + \, \mathcal{O}(\epsilon^4)$ \\

The $N^{(A)}$ term is simple: \\

$N^{(A)} = -J(A, \Psi) = -J(A, \Psi) - \epsilon\widetilde{J}(A, \Psi)$ \\

$N^{(A)} \to -\epsilon^2 J(A_1, \Psi_1) \, - \, \epsilon^3 J(A_1, \Psi_2) \, - \, \epsilon^3 J(A_2, \Psi_1) \, - \, \epsilon^4 J(A_2, \Psi_2) \, - \, \epsilon^3 \widetilde{J}(A_1, \Psi_1) \, - \, \epsilon^4 \widetilde{J}(A_1, \Psi_2) \, - \, \epsilon^4 \widetilde{J}(A_2, \Psi_1) \, - \, \epsilon^5 \widetilde{J}(A_2, \Psi_2)$ \\

And finally, we expand $N^{(B_{1y})} = J(\Psi, B_{1y}) - J(A, u_{1y})$: \\

$= J(\Psi, B_{1y}) + \epsilon\widetilde{J}(\Psi, B_{1y}) - J(A, u_{1y}) - \epsilon\widetilde{J}(A, u_{1y})$ \\

$ = \epsilon^2 J(\Psi_1, B_1) \, + \, \epsilon^3 J(\Psi_1, B_2) \, + \, \epsilon^3 J(\Psi_2, B_1) \, + \, \epsilon^4 J(\Psi_2, B_2) \, + \, \epsilon^3\widetilde{J}(\Psi_1, B_1) \, + \, \epsilon^4\widetilde{J}(\Psi_1, B_2) \, + \, \epsilon^4 \widetilde{J}(\Psi_2, B_1) \, + \, \epsilon^5 \widetilde{J}(\Psi_2, B_2) \, - \, \epsilon^2 J(A_1, u_1) \, - \, \epsilon^3 J(A_1, u_2) \, - \, \epsilon^3 J(A_2, u_1) \, - \, \epsilon^4 J(A_2, u_2) \, - \, \epsilon^3 \widetilde{J}(A_1, u_1) \, - \, \epsilon^4 \widetilde{J}(A_1, u_2) \, - \, \epsilon^4 \widetilde{J}(\Psi_2, B_1) \, - \, \epsilon^5 \widetilde{J}(\Psi_2, B_2)$\\

Thus we are left with the following: \\

$\mathbf{N} = \epsilon^2\mathbf{N_1} \, + \, \epsilon^3\mathbf{N_2} \, + \, \mathcal{O}(\epsilon^4)$ \\

\noindent $N^{(\Psi)} = \epsilon^2 N_2^{(\Psi)} \, + \, \epsilon^3 N_3^{(\Psi)} \, + \, \mathcal{O}(\epsilon^4)$ \\
$N^{(u)} = \epsilon^2 N_2^{(u)} \, + \, \epsilon^3 N_3^{(u)} \, + \, \mathcal{O}(\epsilon^4)$ \\
$N^{(A)} = \epsilon^2 N_2^{(A)} \, + \, \epsilon^3 N_3^{(A)} \, + \, \mathcal{O}(\epsilon^4)$ \\
$N^{(B)} = \epsilon^2 N_2^{(B)} \, + \, \epsilon^3 N_3^{(B)} \, + \, \mathcal{O}(\epsilon^4)$ \\

$N_2^{(\Psi)} = J(\Psi_1, \nabla^2 \Psi_1) \,-\, J(A_1, \nabla^2 A_1) $ \\

$N_2^{(u)} = J(\Psi_1, u_1) \, - \, \frac{1}{4\pi} J(A_1, B_1) $ \\

$N_2^{(A)} = - J(A_1, \Psi_1)$ \\

$N_2^{(B)} =  J(\Psi_1, B_1) \, - \, J(A_1, u_1)$\\

$N_3^{(\Psi)} = J(\Psi_1, \nabla^2\Psi_2) \, - \, J(A_1, \nabla^2 A_2) \,+\, J(\Psi_2, \nabla^2\Psi_1) \, - \, J(A_2, \nabla^2 A_1) \,+\, 2 J(\Psi_1, \partial_z\partial_Z \Psi_1) \, - \, 2 J(A_1, \partial_z\partial_Z A_1) \,+\, \widetilde{J}(\Psi_1, \nabla^2 \Psi_1) \, - \,  \widetilde{J}(A_1, \nabla^2 A_1) $ \\

$N_3^{(u)} = J(\Psi_1, u_2) \, + \, J(\Psi_2, u_1) \, + \, \widetilde{J}(\Psi_1, u_1) \, - \, \frac{1}{4\pi} J(A_1, B_2) \, - \, \frac{1}{4\pi} J(A_2, B_1) \, - \, \frac{1}{4\pi} \widetilde{J}(A_1, B_1)$ \\

$N_3^{(A)} = - J(A_1, \Psi_2) \, - \, J(A_2, \Psi_1) \, - \, \widetilde{J}(A_1, \Psi_1) \,$ \\

$N_3^{(B)} = \, J(\Psi_1, B_2) \, + \, J(\Psi_2, B_1) \, + \, \widetilde{J}(\Psi_1, B_1) \, - \, J(A_1, u_2) \, - \,  J(A_2, u_1) \, - \, \widetilde{J}(A_1, u_1) \, $\\

\section*{Perturbation Series Expansion}

We now expand the fluid quantities $\mathbf{V}$ in a perturbation series $\mathbf{V} = \epsilon\mathbf{V_1} + \epsilon^2\mathbf{V_2} + \epsilon^3\mathbf{V_3} + ...$ \\

We drop the $\partial_t$ terms and our $\mathcal{O}(\epsilon^2)$ equation is: \\

$\epsilon^2 \mathcal{D}\partial_T\mathbf{V} \, + \, \mathbf{N} \, = \, \mathcal{L}\mathbf{V} \, + \, \epsilon\widetilde{\mathcal{L}}_1\partial_Z\mathbf{V} \, + \, \epsilon^2\widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V} \, - \, \epsilon^2\partial_z\mathcal{X}\mathbf{V} \, - \, \epsilon^2\partial_z^3\mathcal{L}_3\mathbf{V} \, + \, \mathcal{O}(\epsilon^3)$ \\

Plugging in our $\mathbf{V}$ expansion, we find: \\

$\epsilon^3 \mathcal{D} \partial_T \mathbf{V_1} \, + \, \epsilon^2\mathbf{N_2} \, + \epsilon^3 \mathbf{N_3} \, = \, \epsilon\mathcal{L}\mathbf{V_1} \, + \, \epsilon^2 \mathcal{L} \mathbf{V_2} \, + \, \epsilon^3 \mathcal{L}\mathbf{V_3} \, + \, \epsilon^2\widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1} \, + \epsilon^3 \widetilde{\mathcal{L}}_1\partial_Z \mathbf{V_2} \, + \, \epsilon^3 \widetilde{\mathcal{L}}_2 \partial_Z^2 \mathbf{V_1} \, - \, \epsilon^3\partial_z \mathcal{X} \mathbf{V_1} \, - \, \epsilon^3\partial_z^3\mathcal{L}_3\mathbf{V_1} \, + \, \mathcal{O}(\epsilon^4)$ \\

To $\mathcal{O}(\epsilon)$, we have: \\

$\mathcal{L}\mathbf{V_1} = 0$ \\

To $\mathcal{O}(\epsilon^2)$: \\

$\mathcal{L}\mathbf{V_2} \, = \, \mathbf{N_2} \, - \, \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1}$ \\

To $\mathcal{O}(\epsilon^3)$: \\

$\mathcal{D}\partial_T \mathbf{V_1} \, + \, \mathbf{N_3} \, = \, \mathcal{L} \mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1\partial_Z\mathbf{V_2} \, + \, \widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V_1} \, - \partial_Z\mathcal{X}\mathbf{V_1} \, - \partial_z^3\mathcal{L}_3\mathbf{V_1}$ \\
